Measuring Individual Partisan Harm in Redistricting

### 2021-09-13 paper/paper.Rmd
finish drafting NJ

---
title: Measuring Individual Partisan Harm in Redistricting
authors:
  - name: Cory McCartan
    thanks: Footnote stuff
    department: Department of Statistics
    affiliation: Harvard University
    location: Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Department of Government
    affiliation: Harvard University
    location: Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - partisan fairness
  - representation
  - redistricting simulation
  - individual standing
bibliography: references.bib
biblio-style: apalike
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
---

@@

```{=html}
<!--
1.  Like PB, focuses on only one part of S-V curve, which may not be relevant
-->
```
## Districting Plan Competitiveness {#sec:competitive}

@@

>>>>

---
title: Measuring Individual Partisan Harm in Redistricting
authors:
  - name: Cory McCartan
    thanks: |
      To whom correspondence should be addressed. 
      The authors thank ____ for helpful comments and discussion.
    department: Department of Statistics
    affiliation: Harvard University
    location: Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Department of Government
    affiliation: Harvard University
    location: Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - partisan fairness
  - representation
  - redistricting simulation
  - individual standing
bibliography: references.bib
biblio-style: apalike
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
---

@@

## Districting Plan Competitiveness {#sec:competitive}

@@

The fairest, best overall utility, and high global, low local utility plans have 8 Democratic and 4 Republican seats at baseline, but the best-utility plan has more extreme vote shares, i.e., less competitive districts, while the high global, low local utility plan has vote shares closer to 50%, with the fairest plan falling somewhere in the middle.

In contrast, the low global, high local utility plan has 9 Democratic seats, 1 Republican seat, and 2 highly competitive seats at baseline.
While generally local utility encourages proportionality and noncompetitive districts, conditional on low global utility (poor proportionality) the local utility measure encourages a Democratic gerrymander which satisfies the state's majority of Democratic voters, at the expense of the minority of Republican voters.

The fair but low local utility plan has a similar vote share pattern to the fair plan, but overall with more competitive districts, which are generally penalized by the local utility measure.
Conversely, the unfair but high local utility plan has 9 Democratic seats, 2 Republican seats, and 1 highly competitive seat at baseline---an unfair Democratic gerrymander---while keeping district vote shares relatively noncompetitive.
